---
title: "An investigation of the predictors of thyroid cancer in patients with thyroid nodules"
author:
  - name: Ovie Edafe
    orcid: 0000-0002-6205-806X
    corresponding: true
    email: ovieedafe@hotmail.co.uk
    roles:
      - Researcher
    affiliations:
      - Department of Oncology & Metabolism, University of Sheffield
  - name: Neil Shephard
    orcid: 000-0001-8301-6857
    corresponding: false
    email:
    roles:
      - Researcher
    affiliations:
      - Research Software Engineer, Department of Computer Science, University of Sheffield
  - name: Karen Sisley
    orcid: 0000-0002-5565-2618
    corresponding: false
    roles:
      - Researcher
    affiliations:
      - Senior Lecturer, Clinical Medicine, School of Medicine and Population Health, University of Sheffield
  - name: Sabapathy P Balasubramanian
    orcid: 0000-0001-5953-2843
    corresponding: false
    roles:
      - Researcher
    affiliations:
      - Directorate of General Surgery, Sheffield Teaching Hospitals NHS Foundation Trust
abstract: |
  An abstract summarising the work undertaken and the overall conclusions can be placed here. Sub-headings are currently
  removed because they conflict with those in the body of the text and mess up the links in the Table of Contents.
keywords:
  - Thyroid nodules
  - Thyroid cancer
plain-language-summary: |
    Plain language summary
key-points:
  - Key point 1
  - Key point 2
date: 2024-04-26
bibliography: references.bib
citation:
  container-title: "Sheffield study on thyroid nodules"
number-sections: true
notebook-preview-options:
  preview: false
execute:
  cache: true
---

{{< include sections/_setup.qmd >}}


## Introduction

Thyroid nodules are common. The challenge in the management of thyroid nodules is differentiating between benign and
malignant nodule thyroid nodules.The use fine needle aspiration and cytology (FNAC) still leaves around 20% of
patients that cannot be classify as benign or malignant. This scenario traditionally leads to diagnostic
hemithyroidectomy for definitive histology. Other clinical variables such as patients' demographics, clinical and
biochemical factors have been shown to be
associated with thyroid cancer in patients with thyroid nodules. This has been utilised in studies evaluating
predictors of thyroid cancer with a view of creating a model to aid prediction.
Standard practice on the management of thyroid nodules does not utilise these non ultrasound and non cytological
factors. Combination of these variables considered to be significant with ultrasound and cytological characteristics may
improve management of patients with thyroid nodules.
Thyroid nodules are increasingly being incidentally detected with increased use of imaging in the evaluation of non
thyroid related pathologies. Thus, leading to increase investigation of thyroid nodules and subsequent increased number
of thyroid operations in non diagnostic cases.
There are morbidities associated with thyroid surgery including scar, recurrent laryngeal nerve injury,
hypothyroidism and hypoparathyroidism.
We performed a systematic review to evaluate for predictors of thyroid cancer specifically in patients presenting
with thyroid nodules.
The systematic review a number of potential important variables that may be useful in the prediction of thyroid cancer
in patients with thyroid nodules. The aim of this study was to evaluate the predictors of thyroid cancer with a view of
improving prediction of thyroid cancer using machine learning techniques.


## Methods

This study was reported as per the Strengthening the Reporting of Observational Studies in Epidemiology (STROBE) guidelines

### Study design

This was a retrospective cohort study.

### Setting

The study was conducted at the Sheffield Teaching hospitals NHS Foundation Trusts. This is a tertiary referral centre
for the management of thyroid cancer

### Participants

We included all consecutive patients who presented with thyroid nodule(s) or that were found to have thyroid nodule(s)
on ultrasound done for thyroid pathology or for other non thyroid related pathologies

### Variables
Variable evaluated was based on findings from a systematic review evaluating predictors of thyroid cancer in patients
with thyroid nodules. Data on the following variables were collected: patient demographics (age, gender, ethnicity),
nodule presentation (incidental nodule, palpable nodule, rapid enlargement, compressive symptoms, vocal paresis), past
medical history (hypertension, Graves' disease, Hashimotos' thyroiditis, family history of thyroid cancer, exposure to
neck radiation), biochemistry (thyroid stimulating hormone, lymphocytes, monocytes), ultrasound characteristics (British
Thyroid Association ultrasound (BTA U), nodule size, solitary nodule, nodule consistency, cervical lymphadenopathy),
Royal College of Pathology (RCP) FNAC  classification, type of thyroid surgery, and histological diagnosis.

### Data source

Data was collected from patients' case notes and electronic patients' database using a standardised
data collection proforma. This was initially piloted on 30 patients  and revised to improve data entry. In addition a
number of variables that were not standard collected during workout of patients were not further checked; these include
body mass index (BMI), serum thyroglobulin, serum triiodothyronine (T3), thyroxine (T4), thyroglobulin antibody (TgAb),
thyroid peroxidase antibody (TP0Ab), and urinary iodine.

### Study size
We sought to have a large data set of at least 100 thyroid nodules with a cancer diagnosis using consecutive sampling
technique.
We aimed for a total of 1500 patients with thyroid nodules to achieve our target sample size. With the use
of modern statistical techniques, we proposed such number will be appropriate to detect important variables if it
exists.

### Data analysis

Data was cleaned and analysed using the R Statistical Software @r_citation and the Tidyverse (@tidyverse),  Tidymodels
(@tidymodels) collection of packages.

### Imputation

The dataset is incomplete and there are missing observations across all variables to varying degrees. In order to
maximise the sample available for analysis imputation was used to infer missing values. The Multivariat Imputation via
Chained Equations ([MICE][mice] and implemented in the eponymous R package @vanBuuren2011Dec) was employed which assumes data is
missing at random (a difficult assumption to formally test). The approach takes each variable with missing data and
attempts to predict it using statistical modelling based on the observed values. In essence it is the same approach as
the statistical methods being employed to try and predict Thyroid Cancer and there are a range of statistical techniques
available which include

### Modelling

We used a selection of statistic modelling techniques to evaluate association between variables and thyroid cancer in
patients with thyroid nodules. The patient population was split into training and testing cohorts in a ratio
of `r train`:`r test` and each model is fitted using the training cohort.
This split ratio is generally used in traditional machine learning techniques. The training set of the
data was used to estimate the relation between variables and thyroid cancer. The larger the training data, the better
it is for the model to learn the trends. The test set was used to determine the accuracy of the model in predicting
thyroid cancer; the bigger the test data the more confidence we have in the model prognostic values. We used simple
randomisation technique for the split to prevent bias in the data split. We ensured that there was no duplicate in the
data sets so any test data was not accidentally trained.
Furthermore, cross validation was used to estimate the accuracy of the various machine learning models.
The k-fold techniques splits the data in ?10 folds, and the data was trained on all but one of the the fold, and the
one fold not trained is used to test the data. This was repeated multiple times using a different fold for test and
the others for training until all the folds is utilised for training and testing. Following multiple training process
with k-fold, we selected the model that has the best predictive value for thyroid cancer in the test cohort.
We also used the leave one out (loo) cross-validation to train and test the data set.In this technique, all but one
observation is use to train the data set and one observation is use to test the data; this is repeated until all the
data test is used for testing and training. The model with the best predictive value was selected.


#### LASSO / Elastic Net

LASSO (Least Absolute Shrinkage and Selection Operatror) and Elastic Net @zou2005 are regression methods that perform variable
selection. The original LASSO method proposed by @tibshirani1996 allows the coefficients for independent/predictor
variables to "shrink" down towards zero, effectively eliminating them from influencing the model, this is often referred
to as L<sub>1</sub> regularisation. The Elastic Net @zou2005 improves on the LASSO by balancing L<sub>1</sub>
regularisation with ridge-regression or L<sub>2</sub> regularisation which helps avoid over-fitting.

Both methods avoid many of the shortcomings/pitfalls of stepwise variable selection @thompson1995 @smith2018 and have
been shown to be more accurate in clinical decision making in small datasets with well code, externally selected
variables @steyerberg2001



#### Random Forest

To add reference
The random forest plot is an extension of the decision tree methodology to reduce variance. Decision trees are very
sensitive to the training data set and can lead to high variance; thus potential issues with generalisation of the
model. The random forest plot selects random observation of the dataset to create multiple decision trees. Random
variables are selected for each tree in the training of the data set. The aggregated output of the generated decision
trees is then used to create an estimate.


#### Gradient Boosting

Gradient boosting is a machine learning algorithm that uses decision tree as a base model. The data is initially trained
on this decision tree, but the initial prediction is weak, thus termed a weak based model. In gradient boosting the process
is iterative; a sequence of decision trees is added to the initial tree. Each tree learns from the prior tree(s) to
improve the model, increasing strength and minimising error.

#### SVM

Support Vector Machines is an approach that allows observation with a binary classifications to be separated using a
hyperplane. It finds a hyperplane that best stratify the two classes i.e benign versus malignant nodules. SVM finds the
hyperplane with the maximum margin of separation between the two classes. The support vectors are the data point that
are positioned close to the margin of the hyperplane and these used to select the most appropraite hyperplane. The
support vectors are the only data points that have an influence on the maximum margin in SVM.

#### Comparision



## Results

```{r}
#| label: n_obs
#| eval: true
#| echo: false
#| warning: false
n_obs <- nrow(df)
```


::: {.panel-tabset}

## Demographics

```{r}
#| label: tbl-patient-demographics
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Demographics of study population"
patient_demo <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("age_at_scan", "gender", "ethnicity")) |>
  gtsummary::tbl_summary() |>
  gtsummary::modify_caption("Demographics of study population")
patient_demo

```

## Clinical Characteristics

```{r}
#| label: tbl-clinical-characteristics
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Clinical characteristics between benign and malignant thyrioid nodules"
clinical_charac <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("age_at_scan",
                  "gender",
                  "incidental_nodule",
                  "palpable_nodule",
                  "rapid_enlargement",
                  "compressive_symptoms",
                  "hypertension",
                  "vocal_cord_paresis",
                  "graves_disease",
                  "hashimotos_thyroiditis",
                  "family_history_thyroid_cancer",
                  "exposure_radiation",
                  "final_pathology",
                  )) |>
  gtsummary::tbl_summary(by = final_pathology) |> add_p() |>
  gtsummary::modify_caption("Clinical characteristics between benign and malignant thyrioid nodules")
clinical_charac

```

## Biomarkers

```{r}
#| label: tbl-biomarkers-variables
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Biochemical variables evaluated between benign and malignant thyroid nodules"
biochem_vars <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("albumin",
                  "tsh_value",
                  "lymphocytes",
                  "monocyte",
                  "final_pathology")) |>
  gtsummary::tbl_summary(by = final_pathology) |>
  gtsummary::modify_caption("Biochemical variables evaluated between benign and malignant thyroid nodules")
biochem_vars

```
## Ultrasound

```{r}
#| label: tbl-ultrasound-characteristics
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Ultrasound characteristics of benign and malignant nodules"
ultrasound_char <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("size_nodule_mm",
                  "solitary_nodule",
                  "bta_u_classification",
                  "consistency_nodule",
                  "cervical_lymphadenopathy",
                  "final_pathology")) |>
  gtsummary::tbl_summary(by = final_pathology) |>
  gtsummary::modify_caption("Ultrasound characteristics of benign and malignant nodules")
ultrasound_char

```

## BTA U
```{r}
#| label: u-class-final-path
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "U class path"
ultrasound_final_path <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("bta_u_classification", "final_pathology")) |>
  gtsummary::tbl_summary(by = bta_u_classification) |>
  gtsummary::modify_caption("BTA U classification by final pathology.")
ultrasound_final_path

```

## Thyroid Classification

```{r}
#| label: thy-class-final-path
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Thyroid classification by Final Pathology"
thy_class_final_path <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("thy_classification", "final_pathology")) |>
  gtsummary::tbl_summary(by = thy_classification) |>
  gtsummary::modify_caption("Thyroid classification by final pathology.")
thy_class_final_path

```

## Cytology

```{r}
#| label: tbl-cytology-characteristics
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Cytological characteristics of benign and malignant nodules"
cytology_char <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("thy_classification",
                  "final_pathology")) |>
  gtsummary::tbl_summary(by = final_pathology) |>
  gtsummary::modify_caption("Cytological characteristics of benign and malignant nodules")
cytology_char
```

:::

@tbl-patient-demographics shows the demographics of patients included in this study. A total of `r n_obs` patients were
included in this study with a median (IQR) age of  `r df$age_at_scan |> stats::quantile(probs=c(0.5))` (
`r df$age_at_scan |> stats::quantile(probs=c(0.25))`-`r df$age_at_scan |> stats::quantile(probs=c(0.75))`).
@tbl-clinical-characteristics shows the distribution of clinical variables evaluated between benign and malignant
thyroid nodules.

### Data Description

Details of data completeness and other descriptive aspects go here.



A summary of the variables that are available in this data set can be found in @tbl-variables.


```{r}
#| label: tbl-data-completeness
#| purl: true
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Overall summary of all variables in the Sheffield dataset."
df_summary <- df |>
  ## NB - We drop the study_id its not a variable, rather its an identifier
  dplyr::select(!(study_id)) |>
  dplyr::ungroup() |>
  gtsummary::tbl_summary(
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c(
      "{N_nonmiss}",
      "{mean} ({sd})",
      "{median} ({p25}, {p75})",
      "{min}, {max}"
    ),
    percent="column",      # Include percentages for categorical variables by "column", "row" or "cell"
    missing="always"           # Exclude missing, options are "no", "ifany" and "always"
  ) |>
  gtsummary::modify_caption("Baseline characteristics and referral source of included patients")
df_summary

```

The completeness of the data is shown in @tbl-data-completeness . Where
variables continuous (e.g. `age` or `size_nodule_mm`) basic summary statistics in the form of mean, standard deviation,
median and inter-quartile range are given. For categorical variables that are logical `TRUE`/`FALSE`
(e.g. `palpable_nodule`) the number of `TRUE` observations and the percentage (of those with observed data for that
variable) are shown along with the number that are _Unknown_. For categorical variables such as `gender` percentages
in each category are reported. For all variables an indication of the number of missing observations is also given and
it is worth noting that there are `r gtsummary::inline_text(df_summary, variable="final_pathology", level="Unknown")`
instances where the `final_pathology` is not known which reduces the sample size to
`r df |> dplyr::filter(!is.na(final_pathology)) |> nrow()`.


#### Missing Data

{{< include sections/_missing.qmd >}}

#### Imputation

{{< include sections/_imputation.qmd >}}

### Modelling

**TODO** - This table feels like duplication of @tbl-data-completeness, perhaps have just one? (`@ns-rse` 2024-07-11).

The predictor variables selected to predict `final_pathology` are shown in @tbl-predictors

```{r}
#| label: tbl-predictors
#| purl: true
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Predictors evaluated in modelling."

df_predictors_evaluated <- df_complete |>
  dplyr::ungroup() |>
  gtsummary::tbl_summary(
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c(
      "{N_nonmiss}",
      "{mean} ({sd})",
      "{median} ({p25}, {p75})",
      "{min}, {max}"
    ),
    percent="column",      # Include percentages for categorical variables by "column", "row" or "cell"
    missing="always"           # Exclude missing, options are "no", "ifany" and "always"
  ) |>
  gtsummary::modify_caption("Variables evaluated in the model")
df_predictors_evaluated
```

<!-- Section that sets up the modelling -->
{{< include sections/_recipe.qmd >}}

The following section is output from a Tidymodel approach to logistic regression to try and work out why variables are
not being included.

```{r}
#| label: logistic-train
#| purl: true
#| eval: true
#| echo: false
#| output: true

## define binary logistic regression model
logistic_model <- logistic_reg() |>
  set_engine("glm")
## add the binary logistic regression model to the thyroid workflow
log_thyroid_workflow <- thyroid_workflow |>
  add_model(logistic_model)
## fit the the workflow to the training data
log_thyroid_fit <- fit(log_thyroid_workflow, data = train)
log_thyroid_fit
## to inspect the fit object
str(log_thyroid_fit)
## use fitted model to make prediction
logistic_model_predictions <- predict(log_thyroid_fit, test) |> bind_cols(test)
## examine the processing steps
log_thyroid_fit |> extract_recipe()
## examine the model
log_thyroid_fit |> extract_fit_parsnip()
## to check if the workflow has been trained
log_thyroid_fit$trained
## I can only see age and gender, unsure why the other variable not present.? excluded due to lack of data ?because they are simply not important ? issue with workflow set up
```



A total of `r df_complete |> nrow()` patients had complete data for the selected predictor variables (see
@tbl-predictors).

Results of the various modelling go here. Each section will show the results along with...

+ LIME/Shaply analysis for explanability of models

#### Logistic Regression

{{< include sections/_logistic.qmd >}}

#### LASSO

{{< include sections/_lasso.qmd >}}

#### Elastic Net

{{< include sections/_elastic.qmd >}}



#### Random Forest
{{< include sections/_rforest.qmd >}}

#### Gradient Boosting

{{< include sections/_gradientboost.qmd >}}


#### SVM

{{< include sections/_svm.qmd >}}


#### Explainability

Which factors are important to classification can be assessed not just by the "importance" but by methods know as
[LIME](https://search.r-project.org/CRAN/refmans/lime/html/lime-package.html) (Local Interpretable Model-Agnostic
Explanations)  @ribeiro2016 and [Shapley
values](https://shap.readthedocs.io/en/latest/example_notebooks/overviews/An%20introduction%20to%20explainable%20AI%20with%20Shapley%20values.html)
@lundberg2017

#### Comparision

Comparing the sensitivity of the different models goes here.

+ Table of sensitivity/specificity/other metrics.
+ ROC curves

## Conclusion

The take-away message is....these things are hard!


## Appendix

### Data Dictionary

```{r}
#| label: tbl-variables
#| purl: true
#| eval: true
#| echo: false
#| warning: false
#| tbl-caption: "Description of variables in the Sheffield Thyroid dataset."
var_labels |>
  as.data.frame() |>
  kable(col.names = c("Description"),
        caption="Description of variables in the Sheffield Thyroid dataset.")
```


[mice]: https://amices.org/mice/
