A summary of the variables that are available in this data set can be found in @tbl-variables.


::: {.panel-tabset}

## Demographics

```{r}
#| label: tbl-patient-demographics
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Demographics of study population"
patient_demo <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("age_at_scan", "gender", "ethnicity")) |>
  gtsummary::tbl_summary() |>
  gtsummary::modify_caption("Demographics of study population")
patient_demo

```

## Clinical Characteristics

```{r}
#| label: tbl-clinical-characteristics
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Clinical characteristics between benign and malignant thyrioid nodules"
clinical_charac <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("age_at_scan",
                  "gender",
                  "incidental_nodule",
                  "palpable_nodule",
                  "rapid_enlargement",
                  "compressive_symptoms",
                  "hypertension",
                  "vocal_cord_paresis",
                  "graves_disease",
                  "hashimotos_thyroiditis",
                  "family_history_thyroid_cancer",
                  "exposure_radiation",
                  "final_pathology",
                  )) |>
  gtsummary::tbl_summary(by = final_pathology) |> add_p() |>
  gtsummary::modify_caption("Clinical characteristics between benign and malignant thyrioid nodules")
clinical_charac

```

## Biomarkers

```{r}
#| label: tbl-biomarkers-variables
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Biochemical variables evaluated between benign and malignant thyroid nodules"
biochem_vars <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("albumin",
                  "tsh_value",
                  "lymphocytes",
                  "monocyte",
                  "final_pathology")) |>
  gtsummary::tbl_summary(by = final_pathology) |>
  gtsummary::modify_caption("Biochemical variables evaluated between benign and malignant thyroid nodules")
biochem_vars

```

## Ultrasound

```{r}
#| label: tbl-ultrasound-characteristics
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Ultrasound characteristics of benign and malignant nodules"
ultrasound_char <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("size_nodule_mm",
                  "solitary_nodule",
                  "bta_u_classification",
                  "consistency_nodule",
                  "cervical_lymphadenopathy",
                  "final_pathology")) |>
  gtsummary::tbl_summary(by = final_pathology) |>
  gtsummary::modify_caption("Ultrasound characteristics of benign and malignant nodules")
ultrasound_char

```

## BTA U

```{r}
#| label: tbl-u-class-final-path
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "U class path"
ultrasound_final_path <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("bta_u_classification", "final_pathology")) |>
  gtsummary::tbl_summary(by = bta_u_classification) |>
  gtsummary::modify_caption("BTA U classification by final pathology.")
ultrasound_final_path

```

## Thyroid Classification

```{r}
#| label: tbl-thy-class-final-path
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Thyroid classification by Final Pathology"
thy_class_final_path <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("thy_classification", "final_pathology")) |>
  gtsummary::tbl_summary(by = thy_classification) |>
  gtsummary::modify_caption("Thyroid classification by final pathology.")
thy_class_final_path

```

## Cytology

```{r}
#| label: tbl-cytology-characteristics
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Cytological characteristics of benign and malignant nodules"
cytology_char <- df |>
  dplyr::ungroup() |>
  dplyr::select(c("thy_classification",
                  "final_pathology")) |>
  gtsummary::tbl_summary(by = final_pathology) |>
  gtsummary::modify_caption("Cytological characteristics of benign and malignant nodules")
cytology_char

```

:::

@tbl-patient-demographics shows the demographics of patients included in this study. A total of `r n_obs` patients were
included in this study with a median (IQR) age of  `r df$age_at_scan |> stats::quantile(probs=c(0.5))` (
`r df$age_at_scan |> stats::quantile(probs=c(0.25))`-`r df$age_at_scan |> stats::quantile(probs=c(0.75))`).
@tbl-clinical-characteristics shows the distribution of clinical variables evaluated between benign and malignant
thyroid nodules.
