The [MICE][mice] package @mice offers a number of different methods for imputing variables (see
[documentation][mice_details]) we have investigated Predictive Mean Matching (PMM), Classification and Regression Trees
(CART) and Random Forests (RF). Four rounds of imputation using each method were made.

A comparison of distributions/proportions before and after imputation are presented below to allow assessment of the
utility of each method.


```{r}
#| label: mice-imputation
#| eval: true
#| echo: false
#| output: false
#| cache: true
## Setup MICE mids for various methods
##

#' Impute missing data and plot the results using the mice package.
#'
#' This is a wrapper around the functionality of the \href{https://amices.org/mice}{mice} and
#' \href{https://amices.org/ggmice}{ggmice} packages for Multivariate Imputation by Chained Equations and visualisation
#' of the resulting imputed datasets. Users should refer to the documentation for details of the options available.
#'
#' The wrapper uses \code{\link[mice]{futuremice}} to perform imputation in parallel using the same number of cores as
#' the requested number of imputations. This speeds up the process but be wary if your computer has limited cores.
#'
#' @param df data.frame|tibble Data frame or tibble of original data.
#' @param imputations int Number of imputations to perform.
#' @param iterations int Number of iterations to perform for imputation.
#' @param method str Method of imputation to use, see the Details section of \code{\link[mice]{mice}}.
#' @param action str Action to take when running \code{\link[mice]{complete}}.
#' @param include bool Logical of whether to include the original dataset in the output.
#' @param seed int Seed for random number generation
#'
#'
impute_data <- function(df = df_complete,
                        outcome_var = "final_pathology",
                        imputations = 4,
                        iterations = 5,
                        method = "pmm",
                        action = "long",
                        include = TRUE,
                        continuous = c("albumin", "tsh_value", "lymphocytes", "monocyte", "size_nodule_mm"),
                        categorical = c("ethnicity",
                                        "incidental_nodule",
                                        "palpable_nodule",
                                        "rapid_enlargement",
                                        "compressive_symptoms",
                                        "hypertension",
                                        "vocal_cord_paresis",
                                        "graves_disease",
                                        "hashimotos_thyroiditis",
                                        "family_history_thyroid_cancer",
                                        "exposure_radiation",
                                        "bta_u_classification",
                                        "solitary_nodule",
                                        "cervical_lymphadenopathy",
                                        "thy_classification"),
                        seed = 123) {
    results <- list()
    ## Setup imputation
    results$mids <- df |>
        dplyr::select(-{{ outcome_var }}) |>
        mice::futuremice(m = imputations,
                         method = method,
                         n.core = imputations,
                         iterations = iterations,
                         parallelseed  = seed)
    ## Generate output dataset,
    results$imputed <- results$mids |>
        mice::complete(action = "long", include = include)
    ## Convert the .imp variable which indicates the imputation set to factor with original dataset labelled as such
    results$imputed <- results$imputed |>
         dplyr::mutate(.imp = factor(.imp,
                                     levels = seq(0, imputations, 1),
                                     labels = c("Original", as.character(seq(1, imputations, 1)))))
    ## We need to bind the outcome variable to each imputed dataset so they can be used in analyses
    outcome = df[[outcome_var]]
    n = imputations
    while(n > 0) {
        outcome = append(outcome, df[[outcome_var]])
        n <- n - 1
    }
    results$imputed <- cbind(results$imputed, outcome)
    colnames(results$imputed) <- stringr::str_replace(colnames(results$imputed), "outcome", outcome_var)
    ## Plot traces of the imputation over iteration
    results$trace <- ggmice::plot_trace(results$mid)
    ## Plot correlation between variables
    results$corr <- ggmice::plot_corr(df,
                                      label = TRUE,
                                      square = TRUE,
                                      rotate = TRUE,
                                      diagonal = FALSE)
    ## Plot histograms of continuous variables
    results$histogram <- list()
    for (var in continuous) {
        results$histogram[[var]] <- ggmice::ggmice(results$mids,
                                                   ggplot2::aes(x = .data[[var]],
                                                                group = .imp)) +
                                        ggplot2::geom_density()
    }
    ## Scatterplots and bar charts for categorical variables
    results$scatter <- list()
    results$bar_chart <- list()
    for (var in categorical) {
        results$scatter[[var]] <- ggmice::ggmice(results$mids,
                                                 ggplot2::aes(x = .imp,
                                                              y = .data[[var]])) +
                                        ggplot2::geom_jitter()
        results$bar_chart[[var]] <- ggmice::ggmice(results$mids,
                                                   ggplot2::aes(x = .data[[var]],
                                                                fill = .imp)) +
                                        ggplot2::geom_bar(position = "dodge")
        ## Above bar chart is often hard to read, instead make a custom bar chart of the propotion in each category by
        ## each imputation
        results$bar_chart_prop[[var]] <- results$imputed |>
                   dplyr::group_by(.imp, .data[[var]]) |>
                   dplyr::summarize(count = n()) |>
                   dplyr::group_by(.imp) |>
                   dplyr::mutate(prop = count / sum(count)) |>
                   ggplot2::ggplot(aes(x = .data[[var]],
                                       y = prop,
                                       fill = .imp)) +
                   ggplot2::geom_bar(stat = "identity", position = "dodge") +
                   ggplot2::theme(legend.position = "none")

    }
    results
}

## Impute using three different methods using the above impute_data() wrapper
imputations = 5
iterations = 5
mice_pmm <- impute_data(method = "pmm",
                        imputations = imputations,
                        iterations = iterations,
                        seed = 684613)
mice_cart <- impute_data(method = "cart",
                        imputations = imputations,
                        iterations = iterations,
                        seed = 1388466)
mice_rf <- impute_data(method = "rf",
                        imputations = imputations,
                        iterations = iterations,
                        seed = 3151358)


```

The distribution of observed (blue) and imputed values for continous variables are shown in the Tab set immediately
below and across all variables the imputed distrubtions follow closely that of the observed indicating that the
imputation methods have worked well for all three methods tested.

::: {.panel-tabset}

## Albumin

```{r}
#| label: fig-mice-imputation-albumin
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$albumin,
                   mice_cart$histogram$albumin,
                   mice_rf$histogram$albumin,
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)
```

## Monocyte

```{r}
#| label: fig-mice-imputation-monocyte
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$monocyte,
                   mice_cart$histogram$monocyte,
                   mice_rf$histogram$monocyte,
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Lymphocytes

```{r}
#| label: fig-mice-imputation-lymphocytes
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$lymphocytes,
                   mice_cart$histogram$lymphocytes,
                   mice_rf$histogram$lymphocytes,
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## TSH Value

```{r}
#| label: fig-mice-imputation-tsh-value
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$tsh_value,
                   mice_cart$histogram$tsh_value,
                   mice_rf$histogram$tsh_value,
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Nodule Size

```{r}
#| label: fig-mice-imputation-size-nodule-mm
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$size_nodule_mm,
                   mice_cart$histogram$size_nodule_mm,
                   mice_rf$histogram$size_nodule_mm,
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

:::

Similarly for the discrete variables the distribution of proportions in the original and each imputed dataset are shown
below. The observed (in pink/peach) always have slightly lower proportions of the observed values because of the
presence of missing (evidences by only one group being present in the `NA` column) but as with the continuous variables
across all imputation methods the proportions are roughly as expected again indicating that imputation has worked well.

**TODO** - Extract the legends from individual plots and add them to the end of each row, see the [cowplot shared
legends article](https://wilkelab.org/cowplot/articles/shared_legends.html) for pointers on how to do this. Should
ideally also get the `fill` colours to align with those used by `ggmice`.

::: {.panel-tabset}

## Incidental Nodule

```{r}
#| label: fig-mice-imputation-incidental-nodule
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$incidental_nodule + theme(legend.position = "None"),
                   ## mice_cart$scatter$incidental_nodule + theme(legend.position = "None"),
                   ## mice_rf$scatter$incidental_nodule + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$incidental_nodule + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$incidental_nodule + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$incidental_nodule + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$incidental_nodule + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$incidental_nodule + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$incidental_nodule + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Palpable Nodule

```{r}
#| label: fig-mice-imputation-palpable-nodule
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$palpable_nodule + theme(legend.position = "None"),
                   ## mice_cart$scatter$palpable_nodule + theme(legend.position = "None"),
                   ## mice_rf$scatter$palpable_nodule + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$palpable_nodule + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$palpable_nodule + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$palpable_nodule + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$palpable_nodule + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$palpable_nodule + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$palpable_nodule + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Rapid Enlargement

```{r}
#| label: fig-mice-imputation-rapid-enlargement
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$rapid_enlargement + theme(legend.position = "None"),
                   ## mice_cart$scatter$rapid_enlargement + theme(legend.position = "None"),
                   ## mice_rf$scatter$rapid_enlargement + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$rapid_enlargement + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$rapid_enlargement + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$rapid_enlargement + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$rapid_enlargement + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$rapid_enlargement + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$rapid_enlargement + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Compressive Symptoms

```{r}
#| label: fig-mice-imputation-compressive-symptoms
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$compressive_symptoms + theme(legend.position = "None"),
                   ## mice_cart$scatter$compressive_symptoms + theme(legend.position = "None"),
                   ## mice_rf$scatter$compressive_symptoms + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$compressive_symptoms + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$compressive_symptoms + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$compressive_symptoms + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$compressive_symptoms + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$compressive_symptoms + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$compressive_symptoms + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Hypertension

```{r}
#| label: fig-mice-imputation-hypertension
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$hypertension + theme(legend.position = "None"),
                   ## mice_cart$scatter$hypertension + theme(legend.position = "None"),
                   ## mice_rf$scatter$hypertension + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$hypertension + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$hypertension + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$hypertension + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$hypertension + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$hypertension + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$hypertension + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Vocal Cord Paresis

```{r}
#| label: fig-mice-imputation-vocal-cord-paresis
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$vocal_cord_paresis + theme(legend.position = "None"),
                   ## mice_cart$scatter$vocal_cord_paresis + theme(legend.position = "None"),
                   ## mice_rf$scatter$vocal_cord_paresis + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$vocal_cord_paresis + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$vocal_cord_paresis + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$vocal_cord_paresis + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$vocal_cord_paresis + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$vocal_cord_paresis + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$vocal_cord_paresis + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Graves Disease

```{r}
#| label: fig-mice-imputation-graves-disease
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$graves_disease + theme(legend.position = "None"),
                   ## mice_cart$scatter$graves_disease + theme(legend.position = "None"),
                   ## mice_rf$scatter$graves_disease + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$graves_disease + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$graves_disease + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$graves_disease + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$graves_disease + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$graves_disease + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$graves_disease + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Hashimotos Thyroiditis

```{r}
#| label: fig-mice-imputation-hashimotos-thyroiditis
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$hashimotos_thyroiditis + theme(legend.position = "None"),
                   ## mice_cart$scatter$hashimotos_thyroiditis + theme(legend.position = "None"),
                   ## mice_rf$scatter$hashimotos_thyroiditis + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$hashimotos_thyroiditis + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$hashimotos_thyroiditis + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$hashimotos_thyroiditis + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$hashimotos_thyroiditis + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$hashimotos_thyroiditis + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$hashimotos_thyroiditis + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Family History

```{r}
#| label: fig-mice-imputation-family-history-thyroid-cancer
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$family_history + theme(legend.position = "None"),
                   ## mice_cart$scatter$family_history + theme(legend.position = "None"),
                   ## mice_rf$scatter$family_history + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$family_history + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$family_history + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$family_history + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$family_history + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$family_history + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$family_history + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Exposure Radiation

```{r}
#| label: fig-mice-imputation-exposure-radiation
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$exposure_radiation + theme(legend.position = "None"),
                   ## mice_cart$scatter$exposure_radiation + theme(legend.position = "None"),
                   ## mice_rf$scatter$exposure_radiation + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$exposure_radiation + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$exposure_radiation + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$exposure_radiation + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$exposure_radiation + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$exposure_radiation + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$exposure_radiation + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Solitary Nodule

```{r}
#| label: fig-mice-imputation-solitary-nodule
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$solitary_nodule + theme(legend.position = "None"),
                   ## mice_cart$scatter$solitary_nodule + theme(legend.position = "None"),
                   ## mice_rf$scatter$solitary_nodule + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$solitary_nodule + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$solitary_nodule + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$solitary_nodule + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$solitary_nodule + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$solitary_nodule + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$solitary_nodule + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## BTA U-Classification

```{r}
#| label: fig-mice-imputation-bta-u-classification
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$bta_u_classification + theme(legend.position = "None"),
                   ## mice_cart$scatter$bta_u_classification + theme(legend.position = "None"),
                   ## mice_rf$scatter$bta_u_classification + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$bta_u_classification + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$bta_u_classification + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$bta_u_classification + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$bta_u_classification + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$bta_u_classification + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$bta_u_classification + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```

## Cervical Lymphadenopathy

```{r}
#| label: fig-mice-imputation-rapid-cervical-lymphadenopathy
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(## mice_pmm$scatter$cervical_lymphadenopathy + theme(legend.position = "None"),
                   ## mice_cart$scatter$cervical_lymphadenopathy + theme(legend.position = "None"),
                   ## mice_rf$scatter$cervical_lymphadenopathy + theme(legend.position = "None"),
                   ## mice_pmm$bar_chart$cervical_lymphadenopathy + theme(legend.position = "None"),
                   ## mice_cart$bar_chart$cervical_lymphadenopathy + theme(legend.position = "None"),
                   ## mice_rf$bar_chart$cervical_lymphadenopathy + theme(legend.position = "None"),
                   mice_pmm$bar_chart_prop$cervical_lymphadenopathy + theme(legend.position = "None"),
                   mice_cart$bar_chart_prop$cervical_lymphadenopathy + theme(legend.position = "None"),
                   mice_rf$bar_chart_prop$cervical_lymphadenopathy + theme(legend.position = "None"),
                   labels = c("PMM", "Cart", "Random Forest"),
                   nrow = 1,
                   ncol = 3)

```



:::


::: {.panel-tabset}

## PMM

```{r}
#| label: tbl-imputation-summary-pmm
#| tbl-caption: Summary of data imputed via PMM.
#| purl: true
#| eval: true
#| echo: false
#| output: true
mice_pmm$imputed |> gtsummary::tbl_summary(by=".imp")
```

## CART

```{r}
#| label: tbl-imputation-summary-cart
#| tbl-caption: Summary of data imputed via CART.
#| purl: true
#| eval: true
#| echo: false
#| output: true
mice_cart$imputed |> gtsummary::tbl_summary(by=".imp")

```

## RF

```{r}
#| label: tbl-imputation-summary-rf
#| tbl-caption: Summary of data imputed via Random Forest.
#| purl: true
#| eval: true
#| echo: false
#| output: true
mice_rf$imputed |> gtsummary::tbl_summary(by = ".imp")
```

:::

[mice]: https://amices.org/mice/
[mice_detail]: https://amices.org/mice/reference/mice.html#details
