The [MICE][mice] package offers a number of different methods for imputing variables (see [documentation][mice_details])
we have investigated Predictive Mean Matching (PMM), Classification and Regression Trees (CART) and Random Forests
(RF). Four rounds of imputation using each method were made.

A comparison of distributions/proportions before and after imputation are presented below to allow assessment of the
utility of each method.


```{r}
#| label: mice-imputation
#| eval: true
#| echo: false
#| output: false
#| cache: true
## Setup MICE mids for various methods
##

#' Impute missing data and plot the results using the mice package.
#'
#' This is a wrapper around the functionality of the \href{https://amices.org/mice}{mice} and
#' \href{https://amices.org/ggmice}{ggmice} packages for Multivariate Imputation by Chained Equations and visualisation
#' of the resulting imputed datasets. Users should refer to the documentation for details of the options available.
#'
#' The wrapper uses \code{\link[mice]{futuremice}} to perform imputation in parallel using the same number of cores as
#' the requested number of imputations. This speeds up the process but be wary if your computer has limited cores.
#'
#' @param df data.frame|tibble Data frame or tibble of original data.
#' @param imputations int Number of imputations to perform.
#' @param iterations int Number of iterations to perform for imputation.
#' @param method str Method of imputation to use, see the Details section of \code{\link[mice]{mice}}.
#' @param action str Action to take when running \code{\link[mice]{complete}}.
#' @param include bool Logical of whether to include the original dataset in the output.
#' @param seed int Seed for random number generation
#'
#'
impute_data <- function(df = df_complete,
                        outcome_var = "final_pathology",
                        imputations = 4,
                        iterations = 5,
                        method = "pmm",
                        action = "long",
                        include = TRUE,
                        continuous = c("albumin", "tsh_value", "lymphocytes", "monocyte", "size_nodule_mm"),
                        categorical = c("ethnicity",
                                        "incidental_nodule",
                                        "palpable_nodule",
                                        "rapid_enlargement",
                                        "compressive_symptoms",
                                        "hypertension",
                                        "vocal_cord_paresis",
                                        "graves_disease",
                                        "hashimotos_thyroiditis",
                                        "family_history_thyroid_cancer",
                                        "exposure_radiation",
                                        "bta_u_classification",
                                        "solitary_nodule",
                                        "cervical_lymphadenopathy",
                                        "thy_classification"),
                        seed = 123) {
    results <- list()
    ## Setup imputation
    results$mids <- df |>
        dplyr::select(-{{ outcome_var }}) |>
        mice::futuremice(m = imputations,
                         method = method,
                         n.core = imputations,
                         iterations = iterations,
                         parallelseed  = seed)
    ## Generate output dataset,
    results$imputed <- results$mids |>
        mice::complete(action = "long", include = include)
    ## Convert the .imp variable which indicates the imputation set to factor with original dataset labelled as such
    results$imputed <- results$imputed |>
         dplyr::mutate(.imp = factor(.imp,
                                     levels = seq(0, imputations, 1),
                                     labels = c("Original", as.character(seq(1, imputations, 1)))))
    ## We need to bind the outcome variable to each imputed dataset so they can be used in analyses
    outcome = df[[outcome_var]]
    n = imputations
    while(n > 0) {
        outcome = append(outcome, df[[outcome_var]])
        n <- n - 1
    }
    results$imputed <- cbind(results$imputed, outcome)
    colnames(results$imputed) <- stringr::str_replace(colnames(results$imputed), "outcome", outcome_var)
    ## Plot traces of the imputation over iteration
    results$trace <- ggmice::plot_trace(results$mid)
    ## Plot correlation between variables
    results$corr <- ggmice::plot_corr(df,
                                      label = TRUE,
                                      square = TRUE,
                                      rotate = TRUE,
                                      diagonal = FALSE)
    ## Plot histograms of continuous variables
    results$histogram <- list()
    for (var in continuous) {
        results$histogram[[var]] <- ggmice::ggmice(results$mids,
                                                   ggplot2::aes(x = .data[[var]],
                                                                group = .imp)) +
                                        ggplot2::geom_density()
    }
    ## Scatterplots and bar charts for categorical variables
    results$scatter <- list()
    results$bar_chart <- list()
    for (var in categorical) {
        results$scatter[[var]] <- ggmice::ggmice(results$mids,
                                                 ggplot2::aes(x = .imp,
                                                              y = .data[[var]])) +
                                        ggplot2::geom_jitter()
        results$bar_chart[[var]] <- ggmice::ggmice(results$mids,
                                                   ggplot2::aes(x = .data[[var]],
                                                                fill = .imp)) +
                                        ggplot2::geom_bar(position = "dodge")
    }
    results
}

## Impute using three different methods using the above impute_data() wrapper
mice_pmm <- impute_data(method = "pmm",
                        imputations = 5,
                        iterations = 5,
                        seed = 684613)
mice_cart <- impute_data(method = "cart",
                        imputations = 5,
                        iterations = 5,
                        seed = 1388466)
mice_rf <- impute_data(method = "rf",
                        imputations = 5,
                        iterations = 5,
                        seed = 3151358)


```

The convergence of the imputation methods are shown in figues @fig-mice-convergence-pmm, @fig-mice-convergence-cart, and
@fig-mice-convergence-rf.

::: {.panel-tabset}

```{r}
#| label: fig-mice-convergence-pmm
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
mice_pmm$trace

```

```{r}
#| label: fig-mice-convergence-cart
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"

mice_cart$trace

```

```{r}
#| label: fig-mice-convergence-rf
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
mice_rf$trace

```

:::


::: {.panel-tabset}

## Albumin

```{r}
#| label: fig-mice-imputation-albumin
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$albumin,
                   mice_cart$histogram$albumin,
                   mice_rf$histogram$albumin,
                   nrow = 1,
                   ncol = 3)
```

## Monocyte

```{r}
#| label: fig-mice-imputation-monocyte
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$monocyte,
                   mice_cart$histogram$monocyte,
                   mice_rf$histogram$monocyte,
                   nrow = 1,
                   ncol = 3)

```

## Lymphocytes

```{r}
#| label: fig-mice-imputation-lymphocytes
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$lymphocytes,
                   mice_cart$histogram$lymphocytes,
                   mice_rf$histogram$lymphocytes,
                   nrow = 1,
                   ncol = 3)

```

## TSH Value

```{r}
#| label: fig-mice-imputation-tsh-value
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$tsh_value,
                   mice_cart$histogram$tsh_value,
                   mice_rf$histogram$tsh_value,
                   nrow = 1,
                   ncol = 3)

```

## Nodule Size

```{r}
#| label: fig-mice-imputation-size-nodule-mm
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$histogram$size_nodule_mm,
                   mice_cart$histogram$size_nodule_mm,
                   mice_rf$histogram$size_nodule_mm,
                   nrow = 1,
                   ncol = 3)

```

:::

**TODO** - Extract the legends from individual plots and add them to the end of each row, see the [cowplot shared
legends article](https://wilkelab.org/cowplot/articles/shared_legends.html) for pointers on how to do this. Should
ideally also get the `fill` colours to align with those used by `ggmice`.

::: {.panel-tabset}

## Incidental Nodule

```{r}
#| label: fig-mice-imputation-incidental-nodule
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$incidental_nodule + theme(legend.position = "None"),
                   mice_cart$scatter$incidental_nodule + theme(legend.position = "None"),
                   mice_rf$scatter$incidental_nodule + theme(legend.position = "None"),
                   mice_pmm$bar_chart$incidental_nodule + theme(legend.position = "None"),
                   mice_cart$bar_chart$incidental_nodule + theme(legend.position = "None"),
                   mice_rf$bar_chart$incidental_nodule + theme(legend.position = "None"),
                   nrow = 2,
                   ncol = 3)

```

## Palpable Nodule

```{r}
#| label: fig-mice-imputation-palpable-nodule
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$palpable_nodule,
                   mice_cart$scatter$palpable_nodule,
                   mice_rf$scatter$palpable_nodule,
                   mice_pmm$bar_chart$palpable_nodule,
                   mice_cart$bar_chart$palpable_nodule,
                   mice_rf$bar_chart$palpable_nodule,
                   nrow = 2,
                   ncol = 3)

```

## Rapid Enlargement

```{r}
#| label: fig-mice-imputation-rapid-enlargement
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$rapid_enlargement,
                   mice_cart$scatter$rapid_enlargement,
                   mice_rf$scatter$rapid_enlargement,
                   mice_pmm$bar_chart$rapid_enlargement,
                   mice_cart$bar_chart$rapid_enlargement,
                   mice_rf$bar_chart$rapid_enlargement,
                   nrow = 2,
                   ncol = 3)

```

## Compressive Symptoms

```{r}
#| label: fig-mice-imputation-compressive-symptoms
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$compressive_symptoms,
                   mice_cart$scatter$compressive_symptoms,
                   mice_rf$scatter$compressive_symptoms,
                   mice_pmm$bar_chart$compressive_symptoms,
                   mice_cart$bar_chart$compressive_symptoms,
                   mice_rf$bar_chart$compressive_symptoms,
                   nrow = 2,
                   ncol = 3)

```

## Hypertension

```{r}
#| label: fig-mice-imputation-hypertension
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$hypertension,
                   mice_cart$scatter$hypertension,
                   mice_rf$scatter$hypertension,
                   mice_pmm$bar_chart$hypertension,
                   mice_cart$bar_chart$hypertension,
                   mice_rf$bar_chart$hypertension,
                   nrow = 2,
                   ncol = 3)

```

## Vocal Cord Paresis

```{r}
#| label: fig-mice-imputation-vocal-cord-paresis
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$vocal_cord_paresis,
                   mice_cart$scatter$vocal_cord_paresis,
                   mice_rf$scatter$vocal_cord_paresis,
                   mice_pmm$bar_chart$vocal_cord_paresis,
                   mice_cart$bar_chart$vocal_cord_paresis,
                   mice_rf$bar_chart$vocal_cord_paresis,
                   nrow = 2,
                   ncol = 3)

```

## Graves Disease

```{r}
#| label: fig-mice-imputation-graves-disease
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$graves_disease,
                   mice_cart$scatter$graves_disease,
                   mice_rf$scatter$graves_disease,
                   mice_pmm$bar_chart$graves_disease,
                   mice_cart$bar_chart$graves_disease,
                   mice_rf$bar_chart$graves_disease,
                   nrow = 2,
                   ncol = 3)

```

## Hashimotos Thyroiditis

```{r}
#| label: fig-mice-imputation-hashimotos-thyroiditis
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$hashimotos_thyroiditis,
                   mice_cart$scatter$hashimotos_thyroiditis,
                   mice_rf$scatter$hashimotos_thyroiditis,
                   mice_pmm$bar_chart$hashimotos_thyroiditis,
                   mice_cart$bar_chart$hashimotos_thyroiditis,
                   mice_rf$bar_chart$hashimotos_thyroiditis,
                   nrow = 2,
                   ncol = 3)

```

## Family History

```{r}
#| label: fig-mice-imputation-family-history-thyroid-cancer
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$family_history,
                   mice_cart$scatter$family_history,
                   mice_rf$scatter$family_history,
                   mice_pmm$bar_chart$family_history,
                   mice_cart$bar_chart$family_history,
                   mice_rf$bar_chart$family_history,
                   nrow = 2,
                   ncol = 3)

```

## Exposure Radiation

```{r}
#| label: fig-mice-imputation-exposure-radiation
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$exposure_radiation,
                   mice_cart$scatter$exposure_radiation,
                   mice_rf$scatter$exposure_radiation,
                   mice_pmm$bar_chart$exposure_radiation,
                   mice_cart$bar_chart$exposure_radiation,
                   mice_rf$bar_chart$exposure_radiation,
                   nrow = 2,
                   ncol = 3)

```

## Solitary Nodule

```{r}
#| label: fig-mice-imputation-solitary-nodule
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$solitary_nodule,
                   mice_cart$scatter$solitary_nodule,
                   mice_rf$scatter$solitary_nodule,
                   mice_pmm$bar_chart$solitary_nodule,
                   mice_cart$bar_chart$solitary_nodule,
                   mice_rf$bar_chart$solitary_nodule,
                   nrow = 2,
                   ncol = 3)

```

## BTA U-Classification

```{r}
#| label: fig-mice-imputation-bta-u-classification
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$bta_u_classification,
                   mice_cart$scatter$bta_u_classification,
                   mice_rf$scatter$bta_u_classification,
                   mice_pmm$bar_chart$bta_u_classification,
                   mice_cart$bar_chart$bta_u_classification,
                   mice_rf$bar_chart$bta_u_classification,
                   nrow = 2,
                   ncol = 3)

```

## Cervical Lymphadenopathy

```{r}
#| label: fig-mice-imputation-rapid-cervical-lymphadenopathy
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
cowplot::plot_grid(mice_pmm$scatter$cervical_lymphadenopathy,
                   mice_cart$scatter$cervical_lymphadenopathy,
                   mice_rf$scatter$cervical_lymphadenopathy,
                   mice_pmm$bar_chart$cervical_lymphadenopathy,
                   mice_cart$bar_chart$cervical_lymphadenopathy,
                   mice_rf$bar_chart$cervical_lymphadenopathy,
                   nrow = 2,
                   ncol = 3)

```



:::


::: {.panel-tabset}

## PMM

```{r}
#| label: tbl-imputation-summary-pmm
#| tbl-caption: Summary of data imputed via PMM.
#| purl: true
#| eval: true
#| echo: false
#| output: true
mice_pmm$imputed |> gtsummary::tbl_summary(by=".imp")
```

## CART

```{r}
#| label: tbl-imputation-summary-cart
#| tbl-caption: Summary of data imputed via CART.
#| purl: true
#| eval: true
#| echo: false
#| output: true
mice_cart$imputed |> gtsummary::tbl_summary(by=".imp")

```

## RF

```{r}
#| label: tbl-imputation-summary-rf
#| tbl-caption: Summary of data imputed via Random Forest.
#| purl: true
#| eval: true
#| echo: false
#| output: true
mice_rf$imputed |> gtsummary::tbl_summary(by = ".imp")
```

:::

[mice]: https://amices.org/mice/
[mice_detail]: https://amices.org/mice/reference/mice.html#details
