
```{r}
### visualising missing data in the cohort

library(naniar)

## visualise missing data for each variable on df
vis_miss(df, cluster = TRUE)

## absolute number of missing data
n_miss(df)

## proportion of missing data in df
prop_miss(df)

## visualising concurrence of missingess combination
gg_miss_upset(df)

## visualise missingness by variable
gg_miss_var(df)
```


```{r}
#| label: tbl-data-completeness
#| purl: true
#| eval: false
#| echo: false
#| warning: false
#| tbl-caption: "Overall summary of all variables in the Sheffield dataset."
df_summary <- df |>
  ## NB - We drop the study_id its not a variable, rather its an identifier
  dplyr::select(!(study_id)) |>
  dplyr::ungroup() |>
  gtsummary::tbl_summary(
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c(
      "{N_nonmiss}",
      "{mean} ({sd})",
      "{median} ({p25}, {p75})",
      "{min}, {max}"
    ),
    percent="column",      # Include percentages for categorical variables by "column", "row" or "cell"
    missing="always"           # Exclude missing, options are "no", "ifany" and "always"
  ) |>
  gtsummary::modify_caption("Baseline characteristics and referral source of included patients")
df_summary

```

The completeness of the original data is shown in tables @tbl-imputation-summary-pmm, @tbl-imputation-summary-cart,
@tbl-imputation-summary-rf, along with summaries from four rounds of imputation for each of three imputation methods.
Where variables continuous (e.g. `age` or `size_nodule_mm`) basic summary statistics in the form of mean, standard
deviation, median and inter-quartile range are given. For categorical variables that are logical `TRUE`/`FALSE`
(e.g. `palpable_nodule`) the number of `TRUE` observations and the percentage (of those with observed data for that
variable) are shown along with the number that are _Unknown_. For categorical variables such as `gender` percentages
in each category are reported. For all variables an indication of the number of missing observations is also given and
it is worth noting that there are `r df |> dplyr::filter(is.na(final_pathology)) |> nrow()`
instances where the `final_pathology` is not known which reduces the sample size to
`r df |> dplyr::filter(!is.na(final_pathology)) |> nrow()`.

More detailed tabulations of missing data by variable are shown in @tbl-naniar-miss-var-summary which shows the number
and percentage of missing data for each variable and by case in @tbl-naniar-miss-case-table which shows how much missing
data each case has. A visualisation of this is shown in @fig-visdat-vis-missing .

**NB** - Currently there is a [bug in the stable release of
Quarto](https://github.com/quarto-dev/quarto-cli/issues/10196) which prevents rendering of the missing data figures. It
is fixed in development version [`v1.6.1`](https://github.com/quarto-dev/quarto-cli/releases/tag/v1.6.1) (currently
available as pre-release, so if things don't render upgrade).

::: {.panel-tabset}

## Variables

```{r}
#| label: tbl-naniar-miss-var-summary
#| tbl-caption: Summary of missing data by variable.
#| purl: true
#| eval: true
#| echo: false
#| output: true
naniar::miss_var_summary(df_complete) |>
  knitr::kable(col.names=c("Variable", "N", "%"),
               caption="Summary of missing data by variable.")
```

## Observations

```{r}
#| label: tbl-naniar-miss-case-table
#| tbl-caption: Summary of missing data by case, how much missing data is there per person?
#| purl: true
#| eval: true
#| echo: false
#| output: true
naniar::miss_case_table(df_complete) |>
  knitr::kable(col.names=c("Missing Variables", "N", "%"),
               caption="Summary of missing data by case, how much missing data is there per person?")
```

:::

```{r}
#| label: fig-visdat-vis-missing
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
visdat::vis_miss(df_complete)
```

The [MICE][mice] package also provides tools for visualising missing data and these are shown in figures
@fig-mice-vis-missing-clinical, @fig-mice-vis-missing-biomarker and @fig-mice-vis-missing-ultrasound.

The columns of these plots, labelled along the top, show the variable, if a cell is blue it indicates data is present,
if it is red it indicates there is missing data. The left-hand side shows the total number of observations for that rows
particular combination of variables with number of missing variables indicated on the right. The first row shows that
for these variables there are 604 observations with zero missing data across the listed variables, the second row
indicates there are 166 observations with _just_ `family_history_thyroid_cancer` but there are some with this missing
_and other variables_. The numbers on the bottom of the figure indicate the total number of missing observations for
that variable (e.g. for `family_history_thyroid_cancer` there is a total of
`r dplyr::filter(df_complete, is.na(family_history_thyroid_cancer)) |> nrow()`  missing observations).

**TODO** - Workout why `out-width: "80%"` isn't applied to these figures and/or how to make the `All` figure readable.

::: {.panel-tabset}

## Clinical

```{r}
#| label: fig-mice-vis-missing-clinical
#| fig-caption: Patterns of missing data for clinical variables.
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
mice_missing_clinical <- df_complete |>
  dplyr::select(final_pathology,
                age_at_scan,
                gender,
                incidental_nodule,
                palpable_nodule,
                rapid_enlargement,
                compressive_symptoms,
                hashimotos_thyroiditis,
                family_history_thyroid_cancer,
                cervical_lymphadenopathy) |>
  mice::md.pattern(rotate.names = TRUE)
```

## Biomarkers

```{r}
#| label: fig-mice-vis-missing-biomarkers
#| fig-caption: Patterns of missing data for biomarkers.
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
mice_missing_biomarkers <- df_complete |>
  dplyr::select(final_pathology,
                age_at_scan,
                gender,
                tsh_value,
                albumin,
                lymphocytes,
                monocyte) |>
  mice::md.pattern(rotate.names = TRUE)
```

## Ultrasound

```{r}
#| label: fig-mice-vis-missing-ultrasound
#| fig-caption: Patterns of missing data for ultrasound.
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
mice_missing_ultrasound <- df_complete |>
  dplyr::select(final_pathology,
                age_at_scan,
                gender,
                bta_u_classification,
                thy_classification,
                size_nodule_mm,
                solitary_nodule) |>
  mice::md.pattern(rotate.names = TRUE)
```

## All

```{r}
#| label: fig-mice-vis-missing-all
#| fig-caption: Patterns of missing data across all variables.
#| purl: true
#| eval: true
#| echo: false
#| output: true
#| out-width: "80%"
mice_missing_ultrasound <- df_complete |>
  dplyr::select(final_pathology,
                age_at_scan,
                gender,
                incidental_nodule,
                palpable_nodule,
                rapid_enlargement,
                compressive_symptoms,
                hashimotos_thyroiditis,
                family_history_thyroid_cancer,
                cervical_lymphadenopathy,
                tsh_value,
                albumin,
                lymphocytes,
                monocyte,
                bta_u_classification,
                thy_classification,
                size_nodule_mm,
                solitary_nodule) |>
  mice::md.pattern(rotate.names = TRUE)
```

:::

[mice]: https://amices.org/mice/
